---
title: "Raport"
author: "Marcin Łukaszyk"
date: "March 31, 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
set.seed(1234)
```

# First Article

Associative Classification in R: arc, arulesCBA, and rCBA.

[Link](https://journal.r-project.org/archive/2019/RJ-2019-048/index.html) to the article.
The peper is from 2018-05-29 so it should be repricleble.
Article contains intorduction to the CBA algorithm.
Our problem is a classification problem.
We want to predict species using the four measurements.
To do it we will use algorithms named before.

First we dived iris data set to a trainig and testting set in 80:20 proportion.
```{r iris}
data("iris")
iris <- iris[sample(seq(nrow(iris))), ]
iris_train <- iris[1:(nrow(iris)*.8), ]
iris_test <- iris[-(1:(nrow(iris)*.8)), ]

kable(head(iris_train))
kable(head(iris_test))
```

As we can se the table differs due to nature of a sample function.
It samples randomly thus making results diffrent each time without setting a seed.
Unfortunately authors didnt set one in the paper or if they did they set diffrent one that is in the soruce code.
We eill use set from source code



## arulesCBA

```{r warning=FALSE, include=FALSE}
library(arules)
library(arulesCBA)
```
```{r }
iris_train_disc <- discretizeDF.supervised(Species ~ ., data = iris_train,
method = "mdlp")
kable(head(iris_train_disc))
```

We got similar resluts.
Thats sign that the algorithm works.

```{r}
trans_train <- as(iris_train_disc, "transactions")
inspect(head(trans_train, n = 3))

```
```{r eval=FALSE}
rules <- apriori(trans_train, parameter = list(support = 0.01, confidence = 0.8),
  appearance = list(rhs = grep("Species=", itemLabels(trans_train), value = TRUE),
  default = "lhs"))
```
```{r}
rules <- mineCARs(Species ~ ., data = trans_train, support = 0.01, confidence = 0.8)
inspect(head(rules,n=3))

```

We come to simmilar yet diffrent result.

## arc

Next we got a "arc" R package.
This is package that implements CBA algorithm.

```{r include=FALSE}
library('arc')
```

We can learn CBA model from iris data set in method shown below.

```{r warning=FALSE}
classifier <- arc::cba(iris_train, "Species")
```
```{r}
inspect(classifier@rules)

```

## arulesCBA

Now we will look at the arulesCBA package.

```{r include=FALSE}
library("arulesCBA")
```

```{r}
classifier <- arulesCBA::CBA(Species ~ ., data = iris_train,
supp = 0.05, confidence = 0.9)
```
```{r}
classifier
```

We got te same results.

## rCBA

Lastly we will look at the rCBA package.

```{r include=TRUE,warning=FALSE} 
library("rCBA")
```
```{r warning=FALSE}
classifier <- rCBA::build(iris_train)
inspect(classifier$model)
```

## Conclusios

This article is repriceable.


# Second Article

In this paper we use "lpirfs" package which is used to analysis of impulse in econometrics.
[Link to the peper](https://journal.r-project.org/archive/2019/RJ-2019-052/index.html)

```{r warning=FALSE,message=FALSE}
library(lpirfs)
library(ggpubr)
library(gridExtra)
library(tidyverse)
```
```{r}
endog_data <- interest_rules_var_data
```
```{r}
results_lin <- lp_lin(endog_data,
                      lags_endog_lin = 4,
                      trend          = 0,
                      shock_type     = 0,
                      confint        = 1.96,
                      hor            = 12)

linear_plots <- plot_lin(results_lin)
```

```{r}
lin_plots_all <- sapply(linear_plots, ggplotGrob)
lp_lin_plots  <- marrangeGrob(lin_plots_all, nrow = ncol(endog_data), 
                                              ncol = ncol(endog_data), top = NULL)

lp_lin_plots
```

We got simmilar resulsts.

```{r}
switching_data <-  if_else(dplyr::lag(endog_data$Infl, 3) > 4.75, 1, 0)
results_nl    <- lp_nl(endog_data,
                        lags_endog_lin  = 4,
                        lags_endog_nl   = 4,
                        trend           = 0,
                        shock_type      = 0,
                        confint         = 1.96,
                        hor             = 12,
                        switching       = switching_data,
                        lag_switching   = F,
                        use_logistic    = F)

 nl_plots <- plot_nl(results_nl)
 
 single_plots <- nl_plots$gg_s1[c(3, 6, 9)]
single_plots[4:6] <- nl_plots$gg_s2[c(3, 6, 9)]
all_plots <- sapply(single_plots, ggplotGrob)
marrangeGrob(all_plots, nrow = 3, ncol = 2, top = NULL)
```

Everything works as intended.

```{r eval=TRUE}

 ag_data       <- ag_data
 sample_start  <- 7
 sample_end    <- dim(ag_data)[1]

 endog_data    <- ag_data[sample_start:sample_end,3:5]

 shock         <- ag_data[sample_start:sample_end, 3]

 results_lin_iv <- lp_lin_iv(endog_data,
                             lags_endog_lin = 4,
                             shock          = shock,
                             trend          = 0,
                             confint        = 1.96,
                             hor            = 20)


 iv_lin_plots    <- plot_lin(results_lin_iv)

 ag_data           <- ag_data
 endog_data        <- ag_data[sample_start:sample_end, 3:5]

 shock             <- ag_data[sample_start:sample_end, 7]

 exog_data         <- ag_data[sample_start:sample_end, 6]

 switching_variable <- ag_data$GDP_MA[sample_start:sample_end] - 0.8

 results_nl_iv <- lp_nl_iv(endog_data,
                           lags_endog_nl     = 3,
                           shock             = shock,
                           exog_data         = exog_data,
                           lags_exog         = 4,
                           trend             = 0,
                           confint           = 1.96,
                           hor               = 20,
                           switching         = switching_variable,
                           use_hp            = FALSE,
                           gamma             = 3)

 plots_nl_iv <- plot_nl(results_nl_iv)

 combine_plots <- list()

 combine_plots[[1]] <- iv_lin_plots[[1]]
 combine_plots[[2]] <- iv_lin_plots[[3]]

 combine_plots[[3]] <- plots_nl_iv$gg_s1[[1]]
 combine_plots[[4]] <- plots_nl_iv$gg_s1[[3]]

 combine_plots[[5]] <- plots_nl_iv$gg_s2[[1]]
 combine_plots[[6]] <- plots_nl_iv$gg_s2[[3]]

 lin_plots_all     <- sapply(combine_plots, ggplotGrob)
 combine_plots_all <- marrangeGrob(lin_plots_all, nrow = 2, ncol = 3, top = NULL)
 combine_plots_all
```

We recreated plots from article.

```{r}
  library(httr)
  library(readxl)
  library(dplyr)
 
  url_jst <-"http://www.macrohistory.net/JST/JSTdatasetR3.xlsx"
  GET(url_jst, write_disk(jst_link <- tempfile(fileext = ".xlsx")))
  jst_data <- read_excel(jst_link, 2L)
 
  jst_data <- jst_data                    %>%
              dplyr::filter(year <= 2013) %>%
              dplyr::select(country, year, everything())
 
  data_set <- jst_data %>%
             mutate(stir    = stir)                         %>%
             mutate(mortgdp = 100*(tmort/gdp))              %>%
             mutate(hpreal  = hpnom/cpi)                    %>%
             group_by(country)                              %>%
             mutate(hpreal  = hpreal/hpreal[year==1990][1]) %>%
             mutate(lhpreal = log(hpreal))                  %>%
   
             mutate(lhpy    = lhpreal - log(rgdppc))        %>%
             mutate(lhpy    = lhpy - lhpy[year == 1990][1]) %>%
             mutate(lhpreal = 100*lhpreal)                  %>%
             mutate(lhpy    = 100*lhpy)                     %>%
             ungroup()                                      %>%
   
             mutate(lrgdp   = 100*log(rgdppc))              %>%
             mutate(lcpi    = 100*log(cpi))                 %>%
             mutate(lriy    = 100*log(iy*rgdppc))           %>%
             mutate(cay     = 100*(ca/gdp))                 %>%
             mutate(tnmort  = tloans - tmort)               %>%
             mutate(nmortgdp = 100*(tnmort/gdp))            %>%
             dplyr::select(country, year, mortgdp, stir, ltrate, 
                                                    lhpy, lrgdp, lcpi, lriy, cay, nmortgdp)
 
 
  data_sample <- seq(1870, 2016)[which(!(seq(1870, 2016) %in%
                                            c(seq(1914, 1918), 
                                            seq(1939, 1947))))]

  results_panel <- lp_lin_panel(data_set  = data_set,  data_sample  = data_sample,
                                endog_data        = "mortgdp", cumul_mult   = TRUE,
                                shock             = "stir",    diff_shock   = TRUE,
                                panel_model       = "within",  panel_effect = "individual",
                                robust_cov        = "vcovSCC", c_exog_data  = "cay",
                                c_fd_exog_data    = colnames(data_set)[c(seq(4,9),11)],
                                l_fd_exog_data    = colnames(data_set)[c(seq(3,9),11)],
                                lags_fd_exog_data = 2,      confint      = 1.67,
                                hor               = 10)
 
  plot_lin_panel <- plot_lin(results_panel)
  plot(plot_lin_panel[[1]])
```

The last plot

```{r}
 results_panel <- lp_nl_panel(data_set          = data_set,  
                              data_sample       = data_sample,
                              endog_data        = "mortgdp", cumul_mult     = TRUE,
                              shock             = "stir",    diff_shock     = TRUE,
                              panel_model       = "within",  panel_effect   = "individual",
                              robust_cov        = "vcovSCC", switching      = "lrgdp",
                              lag_switching     = TRUE,      use_hp         = TRUE,
                              lambda            = 6.25,      gamma          = 10,
                              c_exog_data       = "cay",     
                              c_fd_exog_data    = colnames(data_set)[c(seq(4,9),11)],
                              l_fd_exog_data    = colnames(data_set)[c(seq(3,9),11)], 
                              lags_fd_exog_data = 2,      
                              confint           = 1.67,
                              hor               = 10)
 
  nl_plots      <- plot_nl(results_panel)
  combine_plots <- list(nl_plots$gg_s1[[1]], nl_plots$gg_s2[[1]])
  nonlinear_panel_plots <- marrangeGrob(combine_plots, nrow = 1, ncol = 2, top = NULL)
  nonlinear_panel_plots
```

Whole article is replicable

# Third Article

Last arlicle will be about glmperm packege.
It is used for inferencing with small and moderate sized data sets.
[You can access paper here.](https://journal.r-project.org/archive/2010/RJ-2010-007/index.html)
This packege is not on CRAN so we need to install it manually from github.

```{r}
library(devtools)
install_github('cran/glmperm')
library(glmperm)
```
```{r warning=FALSE}
n <- 20
set.seed(4278)
x1 <- rnorm(n)
x0 <- rnorm(n)+x1
y1 <- ifelse(x0+x1+2*rnorm(n)>0,1,0)
test1 <- prr.test(y1~x0+x1,
var="x0", family=binomial())
x2 <- rbinom(n,1,0.6)
y2 <- ifelse(x1+x2+rnorm(n)>0,1,0)
test2 <- prr.test(y2~x1+x2, var="x1",
nrep=10000,family=binomial())
set.seed(4278)
x1 <- rnorm(n)
x0 <- rnorm(n) + x1
nu <- rgamma(n, shape = 2, scale = 1)
y <- rpois(n, lambda = exp(2) * nu)
test3 <- prr.test(y~x0+x1,
var="x0", family=poisson())
test4 <- prr.test(y~x0, var="x0",
nrep=1000,family=poisson())
```

```{r}
summary(test2)

```

```{r}
library('coin')
data(treepipit, package="coin")
test5<-prr.test(counts~cbpiles+coverstorey
+coniferous+coverregen,data=treepipit,
var="cbpiles",family=poisson())
summary(test5)
```

```{r}
test6<-prr.test(counts~cbpiles+coverstorey
+coniferous+coverregen,data=treepipit,
var="cbpiles",family=quasipoisson())
summary(test6)
```

Everything works.

# Summary

All of given articles are repilicable.
The only problem was were the packege wasn't on cran but it is easily fixed when it is on a git hub page.  
